The \project{unRAVE} catalog

### 2016-09-03 article/unrave.tex
Text updates

The model we introduced in Section~\ref{sec:the-model} is generic in the sense that there 
is no description for how complex the vectorizing function $\vecv(\ell)$ should be, or how
many $K$ labels are used.  Given previous successes using a second-order (quadratic) 
polynomial model \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a 
three-label ($\teff$, $\logg$, [Fe/H]) model for all (main-sequence and giant) stars. Thus, 
for our first set of model experiments,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\end{eqnarray}

@@

\stub{We trained the model initially with no regularization. Could reproduce general structure of a main-sequence and HRD, but the intricacies of the structure was gone. Model appeared to be too simple.  However we can use this model to identify dwarfs/giants,etc. Rather than go to a very complex model, we opted to use two models: one for the main-sequence and one for the giant branch.}

\subsection{A 3-label model for unevolved stars}
\label{sec:the-ms-model}

We constructed a three-label quadratic model for the main-sequence and sub-giant
stars in the \epic\ training set.  In order to set the regularization hyperparameter
$\Lambda$ for this model, we trained 30 models with different regularization strengths,
spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
We then performed leave-one-out cross-validation for each model.  Specifically, for 
each star in the training set: we removed the star; trained the model; and then 
inferred labels from the removed star as if it was a test object. We also performed 
leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
will use as the basis for comparison.  For every model we calculated the bias and
root-mean-square (RMS) deviation between the input labels and those inferred through
cross-validation.

We show the \emph{percentage difference} in the RMS deviation of the labels with respect
to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
and lower envelope represent the boundaries across all labels, showing that the RMS
decreased in \emph{all} labels with increasing regularization.  We found similar
improvements in the bias values, however these were already minimal in the unregularized
case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
we achieve RMS deviations that are about 30\% better than the unregularized model.
Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
At this regularization strength, the bias and RMS values found by leave-one-out
cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
for $\logg$, with $X.XX$~dex and $X.XX$~dex for [Fe/H].

We made no \emph{a priori} judgements on whether a star was likely a giant
or a dwarf star; we inferred labels for all \rave\ spectra using this 
main-sequence model.  The resulting labels for the survey sample are shown
in Figure \ref{fig:hrd-from-individual-models}, where significant structure is 
visible.  The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turnoff
and main-sequence stars relative to the subgiant branch is expected from 
the relative lifetimes of these evolutionary phases.  An overdensity
of stars near the giant branch clump is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.
In Section \ref{sec:joining-the-models} we describe how results from multiple
models are combined to produce a single catalog.

\subsection{A 9-label model for detailed abundances of giant stars}
\label{sec:giant-abundance-model}

@@

We used a quadratic vectorizer for the giant star model, where the quadratic and
cross-terms are expanded in the same way as \ref{eq:vectorizer-three-label}, only
with nine labels.  The regularization hyperparameter $\Lambda$ was set in the same
way described in Section \ref{sec:ms-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
and up to $X$\%.  Here the regularization strength also produces a sparser matrix
of $\vectheta$, with a $\approx20$\% increase in the fraction of zero-valued 
entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
third panel).  Based on the increased model sparsity and decreasing RMS deviation
in the labels, we adopt $\Lambda = X.XX$ for our giant star model.

@@

% All them hax used to combine all the results.

% --> Chi-sq of the sample.
% --> Convex hull? Distance to plane?

@@

\subsection{Jackknifing the training set}
\label{sec:jackknife}

\stub{Should we?}

@@

% We have introduced a non-parametric version of The Cannon that uses strict priors on line formation (censoring masks), and prior probability distributions on astrophysical parameters, using isochrones.
% We ran it on the entire RAVE spectra.
% We deliver effective temperature, surface gravity, and abundances of up to 8 elements for \Nstars.  Our internal and external validation tests suggest that we achieve a typical precision of X~K in $\teff$, 0.0X~dex in $\logg$, and $\approx{}X.XX$ in individual chemical abundances.  This catalog constitutes the largest collection of stellar abundances for stars in the first \project{Gaia} data release.  When combined with positions and 3D velocities from \project{Gaia}, the \project{UNRAVE} catalog will likely be crucial for understanding our local place in the Milky Way.

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and
Sergey Koposov (Cambridge).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

>>>>

The model we introduced in Section~\ref{sec:the-model} is generic in the sense that there 
is no description for how complex the vectorizing function $\vecv(\ell)$ should be, or how
many $K$ labels are used.  Given previous successes using a second-order (quadratic) 
polynomial model \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a 
three-label ($\teff$, $\logg$, [Fe/H]) model for all (main-sequence and giant) stars. Thus, 
for our first set of model experiments the vectorizer $\vecv(\ell_n)$ expands as,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\end{eqnarray}

@@

We used no regularization ($\Lambda = 0$) for this model.  After training the model we
treated all \Nspectra\ spectra as test set objects.  In Figure \ref{fig:3label-test-set}
we show the two-dimensional distribution of effective temperatures $\teff$ and surface 
gravities $\logg$ for all spectra with $\chi_{r}^2 < 3$ and $S/N > 10$.  The main-sequence
and red giant branch are clearly visible.  However, the details of stellar evolution are
no longer present: the sub-giant branch is not discernible, and there are a number of
systematic artefacts (over-densities) in the label space of the test set.  These artefacts
disappear when we require additional quality constraints (e.g., no peculiar morphological
classifications), but the details of the Hertzsprung-Russell diagram are still not
present.  Thus, we concluded that while this model is useful for discerning stellar
classifications (e.g., F2-type giant), the labels are too imprecise.

We chose to adopt a separate model for the main-sequence and the giant branch rather
than switch to a single, higher order model.  This choice allows us to infer
detailed abundance labels for giant stars, and to infer metallicity labels for stars
on the main-sequence and sub-giant branch.  However, adopting two separate models
introduces the challenge of how to combine the results from two models, or how to
assign one star as being a main-sequence/giant star.  In Section~\ref{sec:joining-the-models}
we describe how we will use the 3-label model of the main-sequence and giant branch
(introduced in this section) to discriminate between results from the 3-label 
main-sequence model in Section \ref{sec:unevolved-star-model} and the 9-label giant 
star model in Section \ref{sec:evolved-star-model}.

\subsection{A 3-label model for unevolved stars}
\label{sec:unevolved-star-model}

We constructed a three-label quadratic model using only main-sequence and sub-giant
stars in the \epic\ training set.  In order to set the regularization hyperparameter
$\Lambda$ for this model, we trained 30 models with different regularization strengths,
spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
We then performed leave-one-out cross-validation for each model.  Specifically, for 
each star in the training set: we removed the star; trained the model; and then 
inferred labels from the removed star as if it was a test object. We also performed 
leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
will use as the basis for comparison.  For every model we calculated the bias and
root-mean-square (RMS) deviation between the training set labels and those inferred
through cross-validation.

We show the \emph{percentage difference} in the RMS deviation of the labels with respect
to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
and lower envelope represent the boundaries across all labels, showing that with increasing
regularization, the RMS decreased in \emph{all} labels.  We found similar
improvements in the biases, however these were already minimal in the unregularized
case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
we achieve RMS deviations that are about 30\% better than the unregularized case.
Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
At this regularization strength, the bias and RMS values found by leave-one-out
cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
for $\logg$, with $X.XX$~dex and $X.XX$~dex for [Fe/H].

We inferred labels for all \rave\ spectra using this main-sequence model.  
The resulting labels for the survey sample are shown in Figure \ref{fig:hrd-from-individual-models}.  
The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turn-off
and main-sequence stars relative to the sub-giant branch is expected from 
the relative lifetimes of these evolutionary phases.  An over-density
of stars near the giant branch clump is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.

\subsection{A 9-label model for detailed abundances of giant stars}
\label{sec:evolved-star-model}

@@

We used a quadratic vectorizer for the giant star model, where the quadratic and
cross-terms are expanded in the same way as \ref{eq:vectorizer-three-label}, only
with nine labels.  The regularization hyperparameter $\Lambda$ was set in the same
way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
and up to $X$\%.  Here the regularization strength also produces a sparser matrix
of $\vectheta$, with a $\approx20$\% increase in the fraction of zero-valued 
entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
third panel).  Based on the increased model sparsity and decreasing RMS deviation
in the labels, we adopt $\Lambda = X.XX$ for our giant star model.

@@

% The single model.

@@

@@

We performed an independent analysis of \rave\ spectra using a data-driven approach.
We report effective temperature $\teff$, surface gravity $\logg$, and metallicity [Fe/H]
for X stars on the main-sequence and sub-giant branch.  For another X stars on the red
giant branch, we deliver $\teff$, $\logg$, and a total of X.XX~million elemental abundances:
O, Mg, Al, Ca, Si, Fe, and Ni for X stars.

\stub{Typical precision}

\stub{A key science result that we see already}

\stub{Something about exoplanet host stars, TESS, etc}

This catalog constitutes the largest collection of stellar abundances for stars in the
first \gaia\ data release.  When combined with positions and 3D velocities from \gaia,
we hope the \project{unRAVE} catalog will help advance understanding of the Milky Way's
formation and evolution.

@@

\acknowledgements
We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and Sergey Koposov (Cambridge).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.
